---
title: "AditiBhargavaUCSF"
author: "Swapna Mahurkar-Joshi"
date: "October 12, 2017"
output:
  pdf_document: default
  html_document: default
---
This project includes 1. calculating differences inoutcomes by groups for 3 vairables, interaction; 2.mean and standard deviation for all the 3 outcome measures and 3. calculate sample size

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r, echo = FALSE, include = FALSE}
library(pwr)
library(knitr)
library(reshape)
library(ggplot2)
```

# Import data
```{r}
us <- read.csv("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/otherPIprojects/AditiBhargava_UCSF/data/Ussing chamber data analysis-IBS plasma(1).csv", row.names = 2)
colnames(us)[c(6,8,10,11,12)] <- c("BH_screen","BH_colon","Isc","TER","Flux")
us <- us[,-c(1,5,7,9,13)]
```

# Test normalilty of ditribution
```{r}
hist(us$Isc)
hist(us$TER)
hist(us$Flux)

```
# identify and remove outliers 
```{r}
us <- as.data.frame(us)
# sd3 <- 3*sd(us$Isc)
# us$outlier1 <- ifelse(us$Isc>sd3,1,0)
# sd3 <- 3*sd(abs(us$TER))
# us$outlier2 <- ifelse(abs(us$TER)>sd3,1,0)
lowerq = quantile(us$Isc)[2]
upperq = quantile(us$Isc)[4]
iqr = upperq - lowerq
threshold.upper = (iqr * 3) + upperq; #355.1725
threshold.lower = lowerq - (iqr * 3); #-203.795 
# No outlier found according to these criteria

lowerq = quantile(us$TER)[2]
upperq = quantile(us$TER)[4]
iqr = upperq - lowerq
threshold.upper = (iqr * 3) + upperq; # 24.985
threshold.lower = lowerq - (iqr * 3); # -42.9325
# No outlier found according to these criteria

lowerq = quantile(abs(us$Flux))[2]
upperq = quantile(abs(us$Flux))[4]
iqr = upperq - lowerq
threshold.upper = (iqr *3) + upperq; # 97
threshold.lower = lowerq - (iqr * 3); # -34
# No outlier found according to these criteria

us$Flux <- ifelse(us$Flux>97,NA, us$Flux); #using non-parametric tests, due to limited data points none was excluded
hist(us$Flux)
```

# non-parametric tests
```{r}
library(knitr)
us$dxSexInt <- interaction(us$Dx, us$Gender)

w_p <- function (modelobject) {
     p <- modelobject$p.value
    return(p)
}
us[,c(1:2,4,9)] <- lapply(us[,c(1:2,4,9)],as.factor)

comb1 <- expand.grid(names(us[,c(6:8)]),names(us[,c(1,2)]))

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = us)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=2),3)
colnames(resDf) <- names(us[,c(1,2)])
row.names(resDf) <- names(us[,c(6:8)])
kable(resDf, caption = "Association between IBS plasma variables and  mouse permreability")
```

```{r}
# Differences between C-D;
usCD <- subset(us, us$BH_screen=="C"| us$BH_screen=="D")
usCD$BH_screen <- factor(usCD$BH_screen)
comb1 <- expand.grid(names(usCD[,c(6:8)]),colnames(usCD)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usCD)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usCD)[4]
row.names(resDf) <- names(usCD[,c(6:8)])
kable(resDf, caption = "Association between Bowel Habits C and D and  mouse permreability")

# Differences between M-D;
usMD <- subset(us, us$BH_screen=="M"| us$BH_screen=="D"|us$BH_screen=="U")
usMD$BH_screen <- as.factor(gsub("U","M", usMD$BH_screen))
usMD$BH_screen <- factor(usMD$BH_screen)
comb1 <- expand.grid(names(usMD[,c(6:8)]),colnames(usMD)[4])
models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usMD)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usMD)[4]
row.names(resDf) <- names(usMD[,c(6:8)])
kable(resDf, caption = "Association between Bowel Habits M and D and  mouse permreability")

k_p <- function (modelobject) {
     p <- modelobject$p.value
    return(p)
}

comb1 <- expand.grid(names(us[,c(6:8)]),names(us[,c(4,9)]))

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {kruskal.test(formula = x, data = us)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- as.data.frame(resDf)
resDf$BowelHabits <- round(t(data.frame(lapply(res.models, FUN = k_p))),3)[,1][1:3]
resDf$DxSexInt <- round(t(data.frame(lapply(res.models, FUN = k_p))),3)[,1][4:6]
kable(resDf, caption = "Association between IBS plasma variables and  mouse permreability")

```
# Melting data; Descriptive statistics
```{r}
# Dx
us1 <- us[,-c(3,5)]
mdata <- melt(us1, id=c("Dx","Gender","BH_screen","dxSexInt")) 
DxMean <- cast(mdata, Dx~variable, mean, na.rm = TRUE)
DxMean[,2:4] <- round(DxMean[,2:4],2)
DxSd <- cast(mdata, Dx~variable, sd, na.rm = TRUE)
DxSd[,2:4] <- round(DxSd[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=2, ncol = 4)
row.names(dfStatsPar) <- DxMean[,1]
colnames(dfStatsPar) <- colnames(DxMean)
for ( i in 1:2){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(DxMean[i,j],"(",DxSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric Dx")

# Sex
SexMean <- cast(mdata, Gender~variable, mean, na.rm = TRUE)
SexMean[,2:4] <- round(SexMean[,2:4],2)
SexSd <- cast(mdata, Gender~variable, sd, na.rm = TRUE)
SexSd[,2:4] <- round(SexSd[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=2, ncol = 4)
row.names(dfStatsPar) <- SexMean[,1]
colnames(dfStatsPar) <- colnames(SexMean)
for ( i in 1:2){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(SexMean[i,j],"(",SexSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric Sex")

# BH
BHMean <- cast(mdata, BH_screen~variable, mean, na.rm = TRUE)
BHMean[,2:4] <- round(BHMean[,2:4],2)
BHSd <- cast(mdata, BH_screen~variable, sd, na.rm = TRUE)
BHSd[,2:4] <- round(BHSd[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=5, ncol = 4)
row.names(dfStatsPar) <- BHMean[,1]
colnames(dfStatsPar) <- colnames(BHMean)
for ( i in 1:5){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(BHMean[i,j],"(",BHSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric BH at Screeening")

# Dx-Sex int
IntMean <- cast(mdata, dxSexInt~variable, mean, na.rm = TRUE)
IntMean[,2:4] <- round(IntMean[,2:4],2)
IntSd <- cast(mdata, dxSexInt~variable, sd, na.rm = TRUE)
IntSd[,2:4] <- round(IntSd[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=4, ncol = 4)
row.names(dfStatsPar) <- IntMean[,1]
colnames(dfStatsPar) <- colnames(IntMean)
for ( i in 1:4){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(IntMean[i,j],"(",IntSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric IBS-Sex Interaction")

```

# Power calculations
```{r}
# Isc
## 1. Dx parametric
diff <- abs(73.44 - 85.31)
sdDiff <- sqrt(mean(c(49.29^2,50.17^2)))
ES = diff/sdDiff; 
IscSex <- pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 2. Sex parametric
diff <- abs(92.4 - 62.39)
sdDiff <- sqrt(mean(c(49.77^2,44.72^2)))
ES = diff/sdDiff; 
IscDx <- pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")[1]

## 3. BH parametric (C vs D)
diff <- abs(59.13 - 110.61)
sdDiff <- sqrt(mean(c(57.98^2,35.96^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. BH parametric (M vs D)
diff <- abs(49.17 - 110.61)
sdDiff <- sqrt(mean(c(29.13^2,35.96^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. Interaction parametric (IBS-F vs HC-F)
diff <- abs(78.94 - 119.32)
sdDiff <- sqrt(mean(c(51.96^2,39.09^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
```

```{r}
# TER
## 1. Dx parametric
diff <- abs(9 - 11.93)
sdDiff <- sqrt(mean(c(7.58^2,9.8^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 2. Sex parametric
diff <- abs(11.54 - 8.41)
sdDiff <- sqrt(mean(c(9.87^2,6.36^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 3. BH parametric (C vs D)
diff <- abs(8.31 - 14.83)
sdDiff <- sqrt(mean(c(4.05^2,9.85^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. BH parametric (M vs D)
diff <- abs(2.08 - 14.83)
sdDiff <- sqrt(mean(c(1.78^2,9.85^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. Interaction parametric (IBS-F vs HC-F)
diff <- abs(8.7 - 17.21)
sdDiff <- sqrt(mean(c(8.16^2,12.27^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
```

```{r}
# Flux
## 1. Dx parametric
diff <- abs(28.15 - 23.6)
sdDiff <- sqrt(mean(c(19.89^2,15.96^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 2. Sex parametric
diff <- abs(26.51 - 27.31)
sdDiff <- sqrt(mean(c(10.86^2,24.11^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 3. BH parametric (C vs D)
diff <- abs(20.85 - 44.71)
sdDiff <- sqrt(mean(c(7.06^2,33.81^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. BH parametric (M vs D)
diff <- abs(29.64 - 44.71)
sdDiff <- sqrt(mean(c(10.7^2,33.81^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")

## 4. Interaction parametric (IBS-F vs HC-F)
diff <- abs(23.92 - 42.07)
sdDiff <- sqrt(mean(c(9.22^2,NA^2))) # just one data point on HCF flux
# ES = diff/sdDiff; 
# pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
```

# Plots
```{r, Plots, echo = FALSE}
pList <- list()
for (i in 4:6)
  {
  pList[[i]] <- ggplot(us1, aes(x = us1[,1], y = us1[,i], color = us1[,1])) +   
    geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(us1)[1], y =
                                                               colnames(us1)[i])
 }
 print(pList)
 
 pList <- list()
 
for (i in 4:6)
  {
  pList[[i]] <- ggplot(us1, aes(x = us1[,2], y = us1[,i], color = us1[,2])) +   
    geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(us1)[2], y =
                                                               colnames(us1)[i])
 }
 print(pList)
 
 for (i in 4:6)
  {
  pList[[i]] <- ggplot(us1, aes(x = us1[,3], y = us1[,i], color = us1[,3])) +   
    geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(us1)[3], y =
                                                               colnames(us1)[i])
 }
 print(pList)
 
 for (i in 4:6)
  {
  pList[[i]] <- ggplot(us1, aes(x = us1[,7], y = us1[,i], color = us1[,7])) +   
    geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(us1)[7], y =
                                                               colnames(us1)[i])
 }
 print(pList)

for (i in c(1:3,7))
  for (j in 4:6) {
    ggsave(
      ggplot(us1, aes(x = us1[,i], y = us1[,j], color = us1[,i])) + geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(us1)[i], y = colnames(us1)[j]), file = paste(colnames(us1)[i],paste(colnames(us1)[j],"png", sep = "."),sep = "_"), height = 4,width = 4, units = "in"
           )
  }

```

##########################################################
#Non parametric descriptive statistics # Calculating sample size using these is not recommended.
```{r}
DxMedian <- cast(mdata, Dx~variable, median, na.rm = TRUE)
DxMedian[,2:4] <- round(DxMedian[,2:4],2)
DxIqr <- cast(mdata, Dx~variable, IQR, na.rm = TRUE)
DxIqr[,2:4] <- round(DxIqr[,2:4],2)
dfStatsNonPar <-  matrix(NA, nrow=2, ncol = 4)
row.names(dfStatsNonPar) <- DxMedian[,1]
colnames(dfStatsNonPar) <- colnames(DxMedian)
for ( i in 1:2){
  for (j in 2:4) { 
    dfStatsNonPar[i,j] <- paste(DxMedian[i,j],"(",DxIqr[i,j],")", sep = "")
    }
}
kable(t(dfStatsNonPar), caption = "summary statistics Nonparametric Dx")
```

###############################################################
############# v2 #############################################
# bowel habit X sex interaction
## Females 
```{r}
usF <- subset(us, us$Gender=="F"); dim(usF)

# Differences between IBS HC;
usF$Dx <- factor(usF$Dx)
comb1 <- expand.grid(names(usF[,c(6:8)]),colnames(usF)[1])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usF)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usF)[1]
row.names(resDf) <- names(usF[,c(6:8)])
resDf
kable(resDf, caption = "Association between IBS status in women plasma donors and  mouse permreability")


# Differences between C-D;
usCD <- subset(usF, usF$BH_screen=="C"| usF$BH_screen=="D")
usCD$BH_screen <- factor(usCD$BH_screen)
comb1 <- expand.grid(names(usCD[,c(6:8)]),colnames(usCD)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usCD)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usCD)[4]
row.names(resDf) <- names(usCD[,c(6:8)])
resDf
kable(resDf, caption = "Association of bowel Habits in female plasma donors and  mouse interstinal permreability")

# Differences between C-N;
usCN <- subset(usF, usF$BH_screen=="C"| usF$BH_screen=="N")
usCN$BH_screen <- factor(usCN$BH_screen)
comb1 <- expand.grid(names(usCN[,c(6:8)]),colnames(usCN)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usCN)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usCN)[4]
row.names(resDf) <- names(usCN[,c(6:8)])
resDf
kable(resDf, caption = "Association between Bowel Habits C and N and  mouse permreability within Females")

# Differences between D-N;
usDN <- subset(usF, usF$BH_screen=="D"| usF$BH_screen=="N")
usDN$BH_screen <- factor(usDN$BH_screen)
comb1 <- expand.grid(names(usDN[,c(6:8)]),colnames(usDN)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usDN)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usDN)[4]
row.names(resDf) <- names(usDN[,c(6:8)])
resDf
kable(resDf, caption = "Association between Bowel Habits D and N and  mouse permreability within Females")
```

## Males 
```{r}
usM <- subset(us, us$Gender=="M"); dim(usM)

# Differences between IBS HC;
usM$Dx <- factor(usM$Dx)
comb1 <- expand.grid(names(usM[,c(6:8)]),colnames(usM)[1])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usM)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usM)[1]
row.names(resDf) <- names(usM[,c(6:8)])
kable(resDf, caption = "Association between Dx and  mouse permreability within Males")
resDf

# Differences between C-D;
usCD <- subset(usM, usM$BH_screen=="C"| usM$BH_screen=="D")
usCD$BH_screen <- factor(usCD$BH_screen)
comb1 <- expand.grid(names(usCD[,c(6:8)]),colnames(usCD)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usCD)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usCD)[4]
row.names(resDf) <- names(usCD[,c(6:8)])
kable(resDf, caption = "Association between Bowel Habits C and D and  mouse permreability within Males")
resDf
# Differences between C-N;
usCN <- subset(usM, usM$BH_screen=="C"| usM$BH_screen=="N")
usCN$BH_screen <- factor(usCN$BH_screen)
comb1 <- expand.grid(names(usCN[,c(6:8)]),colnames(usCN)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usCN)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usCN)[4]
row.names(resDf) <- names(usCN[,c(6:8)])
kable(resDf, caption = "Association between Bowel Habits C and N and  mouse permreability within Males")
resDf
# Differences between D-N;
usDN <- subset(usM, usM$BH_screen=="D"| usM$BH_screen=="N")
usDN$BH_screen <- factor(usDN$BH_screen)
comb1 <- expand.grid(names(usDN[,c(6:8)]),colnames(usDN)[4])

models <- lapply(paste(comb1[,1],comb1[,2], sep = "~"), formula)
res.models <- lapply(models, FUN = function(x) {wilcox.test(formula = x, data = usDN)})
# names(res.models) <- paste(comb1[,1],comb1[,2], sep = "~")
resDf <- round(matrix(t(data.frame(lapply(res.models, FUN = w_p))),nrow=3, ncol=1),3)
colnames(resDf) <- colnames(usDN)[4]
row.names(resDf) <- names(usDN[,c(6:8)])
kable(resDf, caption = "Association between Bowel Habits D and N and  mouse permreability within Males")
resDf

```

# Plots interaction
```{r}
### Females
library(ggplot2)
for (i in c(1,4))
  for (j in 6:8) {
    ggsave(
      ggplot(usF, aes(x = usF[,i], y = usF[,j], color = usF[,i])) + geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(usF)[i], y = colnames(usF)[j]), file = paste(colnames(usF)[i],paste(colnames(usF)[j],"Female.png", sep = "."),sep = "_"), height = 4,width = 4, units = "in"
           )
  }

### Males
library(ggplot2)
for (i in c(1,4))
  for (j in 6:8) {
    ggsave(
      ggplot(usM, aes(x = usM[,i], y = usM[,j], color = usM[,i])) + geom_boxplot(outlier.colour = NA) + geom_jitter() + labs(x = colnames(usM)[i], y = colnames(usM)[j]), file = paste(colnames(usM)[i],paste(colnames(usM)[j],"Male.png", sep = "."),sep = "_"), height = 4,width = 4, units = "in"
           )
  }
### 
ggplot(us1, aes(x=dxSexInt,y=Isc,color = dxSexInt)) + geom_boxplot(outlier.shape = NA) +geom_jitter(aes(color = us1$dxSexInt))
ggplot(us1, aes(x=dxSexInt,y=TER,color = dxSexInt)) + geom_boxplot(outlier.shape = NA) +geom_jitter(aes(color = us1$dxSexInt))
ggplot(us1, aes(x=dxSexInt,y=Flux,color = dxSexInt)) + geom_boxplot(outlier.shape = NA) +geom_jitter(aes(color = us1$dxSexInt))
```

```{r}

# BH-Sex int
us1 <- us[,-c(3,5)]
us1$BHSexInt <- interaction(us1$BH_screen,us1$Gender)
us1 <- us1[,c(4,5,6,8)]
mdata <- melt(us1, id="BHSexInt") 
IntMean <- cast(mdata, BHSexInt~variable, mean, na.rm = TRUE)
IntMean[,2:4] <- round(IntMean[,2:4],2)
IntSd <- cast(mdata, BHSexInt~variable, sd, na.rm = TRUE)
IntSd[,2:4] <- round(IntSd[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=8, ncol = 4)
row.names(dfStatsPar) <- IntMean[,1]
colnames(dfStatsPar) <- colnames(IntMean)
for ( i in 1:8){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(IntMean[i,j],"(",IntSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric BH-Sex Interaction")

```
# Sample size based on BH -sex interaction
```{r}

BHSexInt	C.F	          D.F	        M.F	          N.F	          C.M	          D.M	          N.M	        U.M
Isc	      78.98(60.62)	138.35(NA)	49.17(29.13)	119.32(39.09)	29.34(56.04)	101.36(37.77)	51.3(35.97)	44.83(NA)
TER	      -8.52(5.08)	  -22.49(NA)	-2.08(1.78)	  -17.21(12.27)	-8(3.71)	    -12.28(10.32)	-6.64(2.4)	-2.9(NA)
Flux	    19.61(9.68)	  25.4(NA)	  29.64(10.7)	  141.2(107.55)	22.7(0.83)	  227.27(300.92)	17.45(12.44)	12.01(NA)
# Isc
## 1. C.F vs N.F parametric
diff <- abs(78.98 - 119.32)
sdDiff <- sqrt(mean(c(60.62^2,39.09^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 26
## 2. D.F vs N.F parametric
# SD NA 
## 3. M.F vs N.F parametric
diff <- abs(78.98 - 119.32)
sdDiff <- sqrt(mean(c(29.13^2,39.09^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=12

# TER
## 1. C.F vs N.F parametric
diff <- abs(-8.52 - -17.21)
sdDiff <- sqrt(mean(c(5.08^2,12.27^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=19
## 2. D.F vs N.F parametric
# SD NA 
## 3. M.F vs N.F parametric
diff <- abs(-2.08 - -17.21)
sdDiff <- sqrt(mean(c(1.78^2,12.27^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=6

# Flux
## 1. C.F vs N.F parametric
diff <- abs(19.61 - 141.2)
sdDiff <- sqrt(mean(c(9.68^2,107.55^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=7
## 2. D.F vs N.F parametric
# SD NA 
## 3. M.F vs N.F parametric
diff <- abs(29.64 - 141.2)
sdDiff <- sqrt(mean(c(10.7^2,107.55^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=8.4

```
# within men
```{r}

# BHSexInt	C.F	          D.F	        M.F	          N.F	          C.M	          D.M	          N.M	        U.M
# Isc	      78.98(60.62)	138.35(NA)	49.17(29.13)	119.32(39.09)	29.34(56.04)	101.36(37.77)	51.3(35.97)	44.83(NA)
# TER	      -8.52(5.08)	  -22.49(NA)	-2.08(1.78)	  -17.21(12.27)	-8(3.71)	    -12.28(10.32)	-6.64(2.4)	-2.9(NA)
# Flux	    19.61(9.68)	  25.4(NA)	  29.64(10.7)	  141.2(107.55)	22.7(0.83)	  227.27(300.92)	17.45(12.44)	12.01(NA)
# Isc
## 1. C.M vs N.M parametric
diff <- abs(29.34 - 51.3)
sdDiff <- sqrt(mean(c(56.04^2,35.97^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 73
## 2. D.F vs N.F parametric
diff <- abs(101.36 - 51.3)
sdDiff <- sqrt(mean(c(37.77^2,35.97^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=10

# TER
## 1. C.M vs N.M parametric
diff <- abs(-8 - -6.64)
sdDiff <- sqrt(mean(c(3.71^2,2.4^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=83
## 2. D.M vs N.M parametric
diff <- abs(-12.28 - -6.64)
sdDiff <- sqrt(mean(c(10.32^2,2.4^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=29


# Flux
## 1. C.M vs N.M parametric
diff <- abs(22.7 - 17.45)
sdDiff <- sqrt(mean(c(0.83^2,12.44^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=45

## 3. D.M vs N.M parametric
diff <- abs(227.27 - 17.45)
sdDiff <- sqrt(mean(c(300.92^2,12.44^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n=8.4
```
# Dx-Sex int
```{r}
## us1 <- us[,-c(3,5)]
us1$dxSexInt <- interaction(us1$Dx,us1$Gender)
us1 <- us1[,c(4,5,6,7)]
mdata <- melt(us1, id="dxSexInt") 
IntMean <- cast(mdata, dxSexInt~variable, mean, na.rm = TRUE)
IntMean[,2:4] <- round(IntMean[,2:4],2)
IntSd <- cast(mdata, dxSexInt~variable, sd, na.rm = TRUE)
IntSd[,2:4] <- round(IntSd[,2:4],2)

IntVar <- cast(mdata, dxSexInt~variable, var, na.rm = TRUE)
IntVar[,2:4] <- round(IntVar[,2:4],2)

dfStatsPar <-  matrix(NA, nrow=4, ncol = 4)
row.names(dfStatsPar) <- IntMean[,1]
colnames(dfStatsPar) <- colnames(IntMean)
for ( i in 1:4){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(IntMean[i,j],"(",IntSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric Dx-Sex Interaction")

```
# sample size
```{r}

# dxSexInt	HC.F	          IBS.F	          HC.M	        IBS.M
# Isc	      119.32(39.09)	  78.94(51.96)	  51.3(35.97)	  67.93(50.71)
# TER	      -17.21(12.27)	  -8.7(8.16)	    -6.64(2.4)	  -9.29(7.72)
# Flux	    141.2(107.55)	  23.92(9.22)	    17.45(12.44)	123.2(221.88)

# 1. IBS.F vs HC.F parametric

## Isc	
diff <- abs(78.94 - 119.32)
sdDiff <- sqrt(mean(c(51.96^2,39.09^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 21

## TER	
diff <- abs(-8.7 - -17.21)
sdDiff <- sqrt(mean(c(8.16^2,12.27^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 24

## Flux	
diff <- abs(23.92 - 141.2)
sdDiff <- sqrt(mean(c(9.22^2,107.55^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 8

# dxSexInt	HC.F	          IBS.F	          HC.M	        IBS.M
# Isc	      119.32(39.09)	  78.94(51.96)	  51.3(35.97)	  67.93(50.71)
# TER	      -17.21(12.27)	  -8.7(8.16)	    -6.64(2.4)	  -9.29(7.72)
# Flux	    141.2(107.55)	  23.92(9.22)	    17.45(12.44)	123.2(221.88)
## 2. IBS.M vs HC.M parametric
## Isc	
diff <- abs(67.93 - 51.3)
sdDiff <- sqrt(mean(c(50.71^2,35.97^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 110

## TER	
diff <- abs(-9.29 - -6.64)
sdDiff <- sqrt(mean(c(7.72^2,2.4^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 74

## Flux	
diff <- abs(123.2 - 17.45)
sdDiff <- sqrt(mean(c(221.88^2,12.44^2)))
ES = diff/sdDiff; 
pwr.t.test(d = ES, sig.level = 0.05, power = 0.8, type = "two.sample")
# n= 35

```
```{r}
# Interaction p value

summary(lm(us$Isc~ us$Dx * us$Gender))
summary(aov(us1$TER~us1$dxSexInt))
summary(aov(us1$Flux~us1$dxSexInt))
```
```{r}
# HC men-women
# IBS men-women

usHcF  <-  subset(us1, us1$dxSexInt=="HC.F")
usHcM  <-  subset(us1,us1$dxSexInt=="HC.M")
usibsF  <-  subset(us1, us1$dxSexInt=="IBS.F")
usibsM <- subset(us1,us1$dxSexInt=="IBS.M")
usHc_diff <- usHcF[,1:3]-usHcM[,1:3]
usIbs_diff <- usibsF[,1:3]-usibsM[,1:3]
```

```{r}
# BH 

us1 <- us[,c(4,6,7,8)]
mdata <- melt(us1, id="BH_screen") 
IntMean <- cast(mdata, BH_screen~variable, mean, na.rm = TRUE)
IntMean[,2:4] <- round(IntMean[,2:4],2)
IntMean <- IntMean[1:4,]
IntSd <- cast(mdata, BH_screen~variable, sd, na.rm = TRUE)
IntSd[,2:4] <- round(IntSd[,2:4],2)
IntSd <- IntSd[1:4,]
dfStatsPar <-  matrix(NA, nrow=4, ncol = 4)
row.names(dfStatsPar) <- IntMean[,1]
colnames(dfStatsPar) <- colnames(IntMean)
for ( i in 1:4){
  for (j in 2:4) { 
    dfStatsPar[i,j] <- paste(IntMean[i,j],"(",IntSd[i,j],")", sep = "")
    }
}
kable(t(dfStatsPar), caption = "summary statistics parametric BH")
```

