---
title: "Permeability06202016"
author: "Swapna Mahurkar-Joshi"
date: "June 20, 2016"
output: word_document
---
Notes: 

I is for Isc: short circuit current. It represents the secretory properties of the epithelium.

G is the conductance, inverse of the resistance (R). Some people prefer to report G instead of R, although both are indicative of the same thing. It's basically representing the tightness of the epithelial barrier and can depend on the flux of ion. So it's an indication of the paracellular permeability of ions.

FD4 is a small fluorescence dye we use to assess paracellular permeability. We put it on the mucosal side of the biopsy and collect at regular interval from the serosal side. FD4 [ ] is the final concentration of FD4 on the serosal side of the chambers after 2h of experiment. It's representing how much of this molecule went through the mucosa. It's telling us how permeable to small uncharged molecule the epithelium is and is a measure of the leak pathway.

FD4 slope, is the data we get by graphing the concentration of FD4 in the serosal chamber at the different times of collection (every 30 min over 2h). Tells us the speed at which the FD4 goes through the mucosa.

# import permeabilty data

```{r, message=F, warning=F}
# library(phyloseq)
library("ggplot2")
library("scales")
library("grid")
library("ape")
# library(phytools)
# library("dada2")

setwd("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/InetstinalPermeability")
 
per <- read.csv("IBS HC biopsies bilan_122116.csv", sep = ",", row.names = 1); dim(per)
# colnames(per)[c(4,8)] <- c("Sex", "FD4")
per$bowel.habits <- gsub("U","M",per$bowel.habits)

# remove out liers as indicated in the original excel sheet
per$I[per$I>= 200| per$I<= 5] <- NA
per$G[per$G>= 200| per$G<= 5] <- NA
per$R[per$R>= 200| per$R<= 5] <- NA
per$Dx <- ifelse(per$bowel.habits== "C","IBS",ifelse(per$bowel.habits=="D","IBS",ifelse(per$bowel.habits=="M","IBS","HC")))

```

# T-test IBS and healthy controls

```{r, message=F, warning=F}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

write.table(per, file = "permeability_data.txt", sep="\t")
df1 <- as.data.frame(per[,c(2,3,4:9,12,13)])
# df1 <- df1[-18,]
fit1 <- lm(G~Dx*Gender, data = df1)
summary(fit1)
lmp(fit1)
# 0.0004976089

fit1 <- lm(G~Dx, data = df1)
summary(fit1)
lmp(fit1)
#0.235

fit1 <- lm(G~Gender, data = df1)
summary(fit1)
lmp(fit1)
# [1] 0.002172


fun1 <- lapply(1:3,function(i){ y <- paste0('trait',i)
     x <- paste0('Dx',i)
     LHS <- paste(x,'Gender',sep='*')
     form <- as.formula(paste(G, "~",LHS)
                        lm(formula = form, data = df1)
    })
     
     
p.lm.dx <- t(sapply(df1[,c(2:7)], lmp))


ttest.all.tp <- function (x) { z<-t.test(x ~ df1$Dx)$p.value
a <- t.test(x ~ df1$Dx)$statistic
b <- cbind(z,a)
return(b)}

p.df<- t(sapply(df1[,c(2:7)], ttest.all.tp))
colnames(p.df) <- c("p value", "T-statistic")
p.df
# I         0.563683517   0.5872778
# G         0.001129131  -3.6434499
# R         0.020752929   2.6286737
# CCh       0.633469131  -0.4879331
# FD4       0.004565584  -3.0034124
# FD4.slope 0.005347327  -2.9419889
```

Interaction Gender and Dx
```{r}
df1$Interaction1 <- interaction(df1$Dx,df1$Gender)
summary(aov(df1$I~ df1$Interaction1))

summary(aov(df1$G~ df1$Interaction1))
RR# 0.00357

summary(aov(df1$R~ df1$Interaction1))
# 1.28e-05 ***

summary(aov(df1$CCh~ df1$Interaction1))
#0.835

```

# T-test men and women

```{r, message=F, warning=F}

## Men
df.m <- as.data.frame(per[per$Gender=="M",c(2:9,12,13)])
ttest.all.tp <- function (x) { z<-t.test(x ~ df.m$Dx)$p.value
a <- t.test(x ~ df.m$Dx)$statistic
b <- cbind(z,a)
return(b)}
p.df<- t(sapply(df.m[,-c(8,9,10)], ttest.all.tp))
colnames(p.df) <- c("p value", "T-statistic")
p.df
# p value T-statistic
# Age       0.773348848   0.2970628
# I         0.319176688   1.0411188
# G         0.006256202  -3.2421084
# R         0.007100520   3.2537889
# CCh       0.681217331   0.4228133
# FD4       0.059480391  -2.0513352
# FD4.slope 0.024761828  -2.5236776

## women
df.f <- as.data.frame(per[per$Gender=="F",c(2:9,12,13)])
ttest.all.tp <- function (x) { z<-t.test(x ~ df.f$Dx)$p.value
a <- t.test(x ~ df.f$Dx)$statistic
b <- cbind(z,a)
return(b)}

p.df<- t(sapply(df.f[,-c(8,9,10)], ttest.all.tp))
colnames(p.df) <- c("p value", "T-statistic")
p.df
#              p value T-statistic
# Age       0.25659346  -1.1781891
# I         0.61819835   0.5578089
# G         0.06400948  -2.0041367
# R         0.70625691   0.3875993
# CCh       0.57299810  -0.6110097
# FD4       0.05695566  -2.0161671
# FD4.slope 0.07557439  -1.8770971
```

# ANOVA bowel habit

```{r, message=F, warning=F}

df1.ibs <- as.data.frame(per[per$Dx=="IBS",c(2:9,12,13)])
# df1 <- df1[-18,]
aov.all.tp <- function (x) { z<-summary(aov(x ~ df1.ibs$bowel.habits))[[1]][["Pr(>F)"]][[1]]
a <- summary(aov(x ~ df1.ibs$bowel.habits))[[1]][['F value']][[1]]
b <- cbind(z,a)
return(b)}
colnames(p.df) <- c("p value", "F-statistic")
p.df<- t(sapply(df1.ibs[,c(1:7)], aov.all.tp))
p.df
#  [,1]      [,2]
# Age       0.379192952 1.0006862
# I         0.645182431 0.4446862
# G         0.514242579 0.6800240
# R         0.448997643 0.8224958
# CCh       0.002962575 7.0657516
# FD4       0.179160515 1.8184737
# FD4.slope 0.303042803 1.2410639
```
```{r}
# df1.ibs gender differnces (within patients)

ttest.all.tp <- function (x) { z<-t.test(x ~ df1.ibs$Gender)$p.value
a <- t.test(x ~ df1.ibs$Gender)$statistic
b <- cbind(z,a)
return(b)}
p.df<- t(sapply(df1.ibs[,-c(8,9,10)], ttest.all.tp))
colnames(p.df) <- c("p value", "T-statistic")
p.df
# p value T-statistic
# Age       0.594922352  -0.5427329
# I         0.226720732   1.2368225
# G         0.028452401   2.3187659
# R         0.126607432  -1.6002592
# CCh       0.542615084   0.6168435
# FD4       0.006766561   2.8956910
# FD4.slope 0.012931394   2.6395100
> 


df1.hc <- as.data.frame(per[per$Dx=="HC",c(2:9,12,13)])
ttest.all.tp <- function (x) { z<-t.test(x ~ df1.hc$Gender)$p.value
a <- t.test(x ~ df1.hc$Gender)$statistic
b <- cbind(z,a)
return(b)}
p.df<- t(sapply(df1.hc[,-c(8,9,10)], ttest.all.tp))
colnames(p.df) <- c("p value", "T-statistic")
p.df
#               p value T-statistic
# Age       0.258240352  -1.2472137
# I         0.551450327   0.6768987
# G         0.005351960   4.3521725
# R         0.001473566  -4.8112800
# CCh       0.645600377  -0.4987990
# FD4       0.021825862   2.9572880
# FD4.slope 0.044380250   2.4969007

# Controling for gender, the effect goes away
# Dx control for Gender
summary(lm(per$I ~ per$Dx*per$Gender))
# 0.588
summary(lm(per$G ~ per$Dx*per$Gender))
#  0.235
summary(lm(per$R ~ per$Dx*per$Gender))
# 0.82005
summary(lm(per$CCh ~ per$Dx*per$Gender))
# 0.508 
summary(lm(per$FD4 ~ per$Dx*per$Gender))
# 0.19433 
summary(lm(per$FD4.slope ~ per$Dx*per$Gender))
# 0.21218
```


```{r}
# Gender only
summary(lm(per$I ~ per$Gender))
#0.499
summary(lm(per$G ~ per$Gender))
# 0.000471
summary(lm(per$R ~ per$Gender))
# 0.00011
summary(lm(per$CCh ~ per$Gender))
# 0.633
summary(lm(per$FD4 ~ per$Gender))
#  0.00982
summary(lm(per$FD4.slope ~ per$Gender))
# 0.0225
summary(lm(per$I ~ per$Gender))
# 0.499
```

```{r}
# Clinica traits
perMap <- read.csv("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/InetstinalPermeability/sjoshi_permameablity_newsamples_112816.csv", row.names =1)

comPer <- perMap[row.names(perMap)%in%row.names(df1.ibs),];dim(comPer)
df1.ibs.com <- df1.ibs[row.names(df1.ibs)%in%row.names(comPer),]; dim(df1.ibs.com)
comPer <- comPer[row.names(df1.ibs.com),]
match(row.names(comPer), row.names(df1.ibs.com))
df1.ibs.com$AbdominalPain <- comPer$BSQ_AbdPain
df1.ibs.com$Bloating <- comPer$BSQ_Bloating
df1.ibs.com$AbSxIntensity24h <- comPer$AbSxIntensity24h
df1.ibs.com$AbSxIntensity_po_2<- comPer$AbSxIntensity_po_2

summary(lm(df1.ibs.com$I~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.40798

summary(lm(df1.ibs.com$G~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.368

summary(lm(df1.ibs.com$R~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.417

summary(lm(df1.ibs.com$CCh~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.01175*

summary(lm(df1.ibs.com$FD4~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.71895

summary(lm(df1.ibs.com$FD4.slope~df1.ibs.com$AbdominalPain*df1.ibs.com$Gender))
# 0.3357
#########
summary(lm(df1.ibs.com$I~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.97284

summary(lm(df1.ibs.com$G~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.119

summary(lm(df1.ibs.com$R~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.256

summary(lm(df1.ibs.com$CCh~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.0216*

summary(lm(df1.ibs.com$FD4~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.72292

summary(lm(df1.ibs.com$FD4.slope~df1.ibs.com$AbSxIntensity24h*df1.ibs.com$Gender))
# 0.7143

#########
summary(lm(df1.ibs.com$I~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.8733

summary(lm(df1.ibs.com$G~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.398

summary(lm(df1.ibs.com$R~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.727114

summary(lm(df1.ibs.com$CCh~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.0225*

summary(lm(df1.ibs.com$FD4~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.952

summary(lm(df1.ibs.com$FD4.slope~df1.ibs.com$AbSxIntensity_po_2*df1.ibs.com$Gender))
# 0.76795



### bloating

summary(lm(df1.ibs.com$I~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.5165

summary(lm(df1.ibs.com$G~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.3448

summary(lm(df1.ibs.com$R~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.5742

summary(lm(df1.ibs.com$CCh~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.7138

summary(lm(df1.ibs.com$FD4~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.05814

summary(lm(df1.ibs.com$FD4.slope~df1.ibs.com$Bloating*df1.ibs.com$Gender))
# 0.3568

# Bowel habit
per.ibs<- subset(per, per$Dx == "IBS")
summary(lm(per.ibs$I ~ per.ibs$bowel.habits* per.ibs$Gender)) #NS
summary(aov(per.ibs$I ~ per.ibs$bowel.habits))
#0.566

summary(lm(per.ibs$G ~ per.ibs$bowel.habits)) #NS
summary(aov(per.ibs$G ~ per.ibs$bowel.habits))
# 0.86

summary(lm(per.ibs$R ~ per.ibs$bowel.habits)) #NS
summary(aov(per.ibs$R ~ per.ibs$bowel.habits))
# 0.672

summary(lm(per.ibs$CCh ~ per.ibs$bowel.habits* per.ibs$Gender)) 
# CCH is significantly higher in IBS-C compared to IBS D and HC 0.00398, 0.00341
summary(aov(per.ibs$CCh ~ per.ibs$bowel.habits))
# 0.00487

summary(lm(per.ibs$FD4 ~ per.ibs$bowel.habits)) #NS
summary(aov(per.ibs$FD4 ~ per.ibs$bowel.habits))
# 0.238

summary(lm(per.ibs$FD4.slope ~ per.ibs$bowel.habits)) #NS
summary(aov(per.ibs$FD4.slope ~ per.ibs$bowel.habits))
# 0.355

```

# IBS and Control plots


```{r, message=F, warning=F}

setwd("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/InetstinalPermeability/DDW_2016abstract_analysis/")

library(reshape)
library(ggplot2)
library(gridExtra)
library(grid)

# df2 <- melt(df1, id = c("dx","Sex", "bowel.habits","menstrual.cycle"))

p <- ggplot(per, aes(as.factor(Dx),I, fill = per$Dx)) + geom_boxplot() 
ggsave(p, file= "Dx_I.png")

p <- ggplot(per, aes(as.factor(Dx),G, fill = factor(Dx)))
p2 <- p + geom_boxplot()
ggsave(p2, file= "Dx_G.png")

p <- ggplot(per, aes(as.factor(Dx),R, fill = factor(Dx)))
p3 <- p + geom_boxplot()
ggsave(p3, file= "Dx_R.png")

p <- ggplot(per, aes(as.factor(Dx),FD4, fill = factor(Dx)))
p4 <- p + geom_boxplot()
ggsave(p4, file= "Dx_FD4.png")

p <- ggplot(per, aes(as.factor(Dx),FD4.slope, fill = factor(Dx)))
p5 <- p + geom_boxplot()
ggsave(p5, file= "Dx_FD4_slope.png")

p <- ggplot(per, aes(as.factor(Dx),CCh, fill = factor(Dx)))
p6 <- p + geom_boxplot()
ggsave(p6, file= "Dx_CCh.png")

# grid.arrange(p1,p2,p3,p4,p5, ncol=1, nrow = 5)
ggsave(grid.arrange(p1,p2,p3,p4,p5, ncol=2, nrow = 3), file = "IBSpermeability.png", height = 7, width = 10, units = "in")

p1
p2
p3
p4
p5
```

# IBS and gender interaction plots

```{r, message=F, warning=F}
p <- ggplot(per, aes(as.factor(Gender),I, fill = per$Gender)) + geom_boxplot() 
ggsave(p, file= "Gender_I.png")

p <- ggplot(per, aes(as.factor(Gender),G, fill = per$Gender))
p2 <- p + geom_boxplot()
ggsave(p2, file= "Gender_G.png")

p <- ggplot(per, aes(as.factor(Gender),R, fill = factor(Gender)))
p3 <- p + geom_boxplot()
ggsave(p3, file= "Gender_R.png")

p <- ggplot(per, aes(as.factor(Gender),FD4, fill = factor(Gender)))
p4 <- p + geom_boxplot()
ggsave(p4, file= "Gender_FD4.png")

p <- ggplot(per, aes(as.factor(Gender),FD4.slope, fill = factor(Gender)))
p5 <- p + geom_boxplot()
ggsave(p5, file= "Gender_FD4_slope.png")

p <- ggplot(per, aes(as.factor(Gender), CCh, fill = factor(Gender)))
p6 <- p + geom_boxplot()
ggsave(p6, file= "Gender_CCh.png")

per$dxGenInt <- interaction(per$Dx,per$Gender)
p <- ggplot(per, aes(as.factor(dxGenInt),I, fill = factor(dxGenInt)))
p1 <- p + geom_jitter(width = 0.25, size = 1.5)+ ylim (0,150) + geom_errorbar(yintercept = "mean")

p <- ggplot(per, aes(as.factor(dxGenInt),G, fill = factor(dxGenInt)))
p2 <- p + geom_jitter(width = 0.25, size = 1.5)+ ylim (0,150)

p <- ggplot(per, aes(as.factor(dxGenInt),R, fill = factor(dxGenInt)))
p3 <- p + geom_jitter(width = 0.25, size = 1.5) + ylim (0,40)

p <- ggplot(per, aes(as.factor(dxGenInt),FD4, fill = factor(dxGenInt)))
p4 <- p + geom_jitter(width = 0.25, size = 1.5) + ylim (0,500)

p <- ggplot(per, aes(as.factor(dxGenInt),FD4.slope, fill = factor(dxGenInt)))
p5 <- p + geom_jitter(width = 0.25, size = 1.5) + ylim (0,1.5)

# grid.arrange(p1,p2,p3,p4,p5, ncol=2, nrow = 3)
ggsave(grid.arrange(p1,p2,p3,p4,p5, ncol=1, nrow = 5), file = "IBSpermeability_gender_interaction.png",height = 10, width = 8, units = "in")
p2
p3
p4
p5
```

# IBS and gender interaction plots with bowel habits jitter

```{r, message=F, warning=F}

per$bowel.habits <- gsub("U","M", per$bowel.habits)
per$dxGenInt <- interaction(per$Dx,per$Gender)
per$BH_color <- ifelse(per$bowel.habits == "C", "red", ifelse(per$bowel.habits == "D", "Green",ifelse(per$bowel.habits == "M", "purple", "grey")))

p <- ggplot(per, aes(as.factor(dxGenInt),I))
p1 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p1, file = "dxGenInt_I.png")

p <- ggplot(per, aes(as.factor(dxGenInt),G))
p2 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p2, file = "dxGenInt_G.png")


p <- ggplot(per, aes(as.factor(dxGenInt),R))
p3 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p3, file = "dxGenInt_R.png")

p <- ggplot(per, aes(as.factor(dxGenInt),FD4))
p4 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p4, file = "dxGenInt_FD4.png")

p <- ggplot(per, aes(as.factor(dxGenInt),FD4.slope))
p5 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p5, file = "dxGenInt_FD4_slope.png")

p <- ggplot(per, aes(as.factor(dxGenInt),CCh))
p5 <- p + geom_boxplot()+ geom_jitter(size = 3, color = per$BH_color)
ggsave(p5, file = "dxGenInt_CCh.png")
# grid.arrange(p1,p2,p3,p4,p5, ncol=1, nrow = 5)
ggsave(grid.arrange(p1,p2,p3,p4,p5, ncol=1, nrow = 5), file = "IBSpermeability_Dx_BH_Gender.png",height = 7, width = 10, units = "in")
p1
p2
p3
p4
p5
```

# Bowel habit differences plots

```{r, message=F, warning=F}

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),I, fill =bowel.habits))
p1 <- p + geom_boxplot()
ggsave(p1, file= "BH_I.png")

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),G, fill =bowel.habits))
p2 <- p + geom_boxplot()
ggsave(p2, file = "BH_G.png")

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),R, fill =bowel.habits))
p3 <- p + geom_boxplot()
ggsave(p3, file= "BH_R.png")

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),CCh, fill =bowel.habits))
p3 <- p + geom_boxplot()
ggsave(p3, file= "BH_CCh.png")

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),FD4, fill =bowel.habits))
p4 <- p + geom_boxplot()
ggsave(p4, file= "BH_FD4.png")

p <- ggplot(per.ibs, aes(as.factor(bowel.habits),FD4.slope, fill =bowel.habits))
p5 <- p + geom_boxplot()
ggsave(p5, file = "BH_FD4_slope.png")



p1
p2
p3
p4
p5
# grid.arrange(p1,p2,p3,p4,p5, ncol=1, nrow = 5)
ggsave(grid.arrange(p1,p2,p3,p4,p5, ncol=2, nrow = 3), file = "IBSpermeability_BH.png",height = 7, width = 10, units = "in")
```

Interaction with jitter

```{r}
p <- ggplot(per, aes(as.factor(dxGenInt),I, fill = dxGenInt))
p1 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p1, file = "dxGenInt_I.png")

p <- ggplot(per, aes(as.factor(dxGenInt),G,  fill = dxGenInt))
p2 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p2, file = "dxGenInt_G.png")


p <- ggplot(per, aes(as.factor(dxGenInt),R, fill = dxGenInt))
p3 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p3, file = "dxGenInt_R.png")

p <- ggplot(per, aes(as.factor(dxGenInt),FD4,  fill = dxGenInt))
p4 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p4, file = "dxGenInt_FD4.png")

p <- ggplot(per, aes(as.factor(dxGenInt),FD4.slope,  fill = dxGenInt))
p5 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p5, file = "dxGenInt_FD4_slope.png")

p <- ggplot(per, aes(as.factor(dxGenInt),CCh,  fill = dxGenInt))
p5 <- p + geom_boxplot()+ geom_jitter(size = 3)
ggsave(p5, file = "dxGenInt_CCh.png")

```




################################################################### unused for DDW
# dichotomize permeability at the median

```{r}
# select G/conductance as a measure of permeability
perMedian <- median(per$G, na.rm = TRUE) 
perMedian # 67.46633

# divide the samples into two gropus highPerm and lowPerm based on the median

per$highPerm <- ifelse(per$G >= perMedian, 1,0)

fisher.test(table(per$dx,per$highPerm))

save(per, file = "per.rda")

```

```{r}
library(graphics)
png("mosaicplotPerm.png")
mosaicplot (table(per$dx,per$highPerm), color=TRUE) 
dev.off()
mosaicplot (table(per$dx,per$highPerm), color=TRUE) 
```

# fitting logistic regression for trait data with 'G' 

```{r}
# import phenotype data
pheno <- read.csv("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/InetstinalPermeability/sjoshi_permeability_071216.csv", sep = ",", row.names = 1)
row.names(per) <- gsub(" ","",row.names(per))
pheno1 <- pheno[row.names(pheno)%in%row.names(per), ]
per1 <- per[row.names(per)%in%row.names(pheno),]
pheno1 <- pheno1[row.names(per1),]
match(row.names(pheno1),row.names(per1))
perPheno <- cbind(pheno1, per1)

perPheno1<- perPheno[perPheno$Group==2,c(4,5,6,8,17, 18,19,20,21,22,23,24,25,26,27,31,33,35,36,37,38,39,40,41,42,43,72,73,74,75,76, 77,80,81,82,83, 84, 85)]

# regression

# calculate p value and plot the lm results

# perPheno1[,3]<- as.numeric(as.character(perPheno1[,3]))
# perPheno1[,2]<- as.numeric(as.character(perPheno1[,2]))
pval <- matrix(NA, nrow = 38, ncol=2)


for ( i in c(1:38))
{
  pval[i,1] = round(anova(lm(perPheno1[,i]~perPheno1[,35]))$'Pr(>F)'[1],3)
pval[i,2] = round(summary(lm(perPheno1[,i]~perPheno1[,35]))$r.squared,3)
png(paste(colnames(perPheno1)[i],"_perm_G.png", sep = ""))
        plot(perPheno1[,i] ~ perPheno1[,35], xlab = "G", ylab = colnames(perPheno1)[i])
        legend("bottomright",paste(paste("p_value =", pval[i,1], paste("RSquared = ", pval[i,2]))), cex = 0.9)
        # text(paste(paste("p_value =", pval[i,1], paste("RSquared = ", pval[i,2]))), cex=0.6, pos=1, col="red")
        dev.off()
        plot(perPheno1[,i] ~ perPheno1[,35], xlab = "G", ylab = colnames(perPheno1)[i])
        legend("bottomright",paste(paste("p_value =", pval[i,1], paste("RSquared = ", pval[i,2]))), cex = 0.9)
        # text(paste(paste("p_value =", pval[i,1], paste("RSquared = ", pval[i,2]))), cex=0.6, pos=1, col="red")
}


row.names(pval) <- colnames(perPheno1)
colnames(pval) <- c("P_value", "RSquared")
pval

# AbdominalPain = 0.078


subset(pval, pval[,1] < 0.1) 



# Individual box plots for vectors, interaction and sample size


##############
perPheno2 <- subset(perPheno1, perPheno1$G != 'NA')
## gender differences
library(grid)
summary(lm(perPheno2[,35]~perPheno2[,3]))
p_value <- round(anova(lm(perPheno2[,35]~perPheno2[,3]))$'Pr(>F)'[1],3)
p <- ggplot(perPheno2, aes(as.factor(Sex),G, fill = factor(Sex)))
grob <- grobTree(textGrob(paste("p_value =", p_value), x=0.5,  y=0.95,  gp=gpar(col="red", fontsize=14)))
p1 <- p + geom_boxplot() + ylim (0,150)+ annotation_custom(grob)
p1

## IBS differences
summary(lm(perPheno2[,35]~perPheno2[,1]))
p_value <- round(anova(lm(perPheno2[,35]~perPheno2[,1]))$'Pr(>F)'[1],3)
p <- ggplot(perPheno2, aes(as.factor(Group),G, fill = factor(Group)))
grob <- grobTree(textGrob(paste("p_value =", p_value), x=0.5,  y=0.95,  gp=gpar(col="red", fontsize=14)))
p1 <- p + geom_boxplot() + ylim (0,150)+ annotation_custom(grob)
p1

## interaction 
# Boxplots

p <- ggplot(perPheno2, aes(as.factor(dxGenInt),G))
p1 <- p + geom_boxplot()+ geom_jitter(size = 3, color = perPheno2$BH_color)+ ylim (0,150)
p1

summary(lm(perPheno2[,35]~perPheno2[,1]*perPheno2[,3]))
p_value <- round(anova(lm(perPheno2[,35]~perPheno2[,1]*perPheno2[,3]))$'Pr(>F)'[1],3)
p <- ggplot(perPheno2, aes(as.factor(dxGenInt),G, fill = factor(dxGenInt)))
grob <- grobTree(textGrob(paste("p_value =", p_value), x=0.5,  y=0.95,  gp=gpar(col="red", fontsize=14)))
p1 <- p + geom_boxplot() + ylim (0,150)+ annotation_custom(grob)
p1
# which groups are significantly different?
combinations <- combn(names(table(perPheno2$dxGenInt)),2)
#      [,1]    [,2]   [,3]    [,4]    [,5]    [,6]   
# [1,] "HC.F"  "HC.F" "HC.F"  "IBS.F" "IBS.F" "HC.M" 
# [2,] "IBS.F" "HC.M" "IBS.M" "HC.M"  "IBS.M" "IBS.M"

int.p <- matrix(NA, nrow = 6, ncol =1)
row.names(int.p) <- paste(combinations[1,], combinations[2,], sep = "_")

int.p[1,1] <- t.test(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="IBS.F",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="IBS.F",39]) )$p.value

int.p[2,1] <- t.test(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="HC.M",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="HC.M",39]) )$p.value

int.p[3,1] <- t.test(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="IBS.M",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "HC.F"| perPheno2$dxGenInt =="IBS.M",39]) )$p.value

int.p[4,1] <- t.test(perPheno2[perPheno2$dxGenInt == "IBS.F"| perPheno2$dxGenInt =="HC.M",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "IBS.F"| perPheno2$dxGenInt =="HC.M",39]) )$p.value

int.p[5,1] <- t.test(perPheno2[perPheno2$dxGenInt == "IBS.F"| perPheno2$dxGenInt =="IBS.M",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "IBS.F"| perPheno2$dxGenInt =="IBS.M",39]) )$p.value

int.p[6,1] <- t.test(perPheno2[perPheno2$dxGenInt == "HC.M"| perPheno2$dxGenInt =="IBS.M",38] ~ as.character(perPheno2[perPheno2$dxGenInt == "HC.M"| perPheno2$dxGenInt =="IBS.M",39]) )$p.value


int.p

# interaction Plots
#1
# library(stats)
# interaction.plot(as.factor(perPheno2$Group),as.factor(perPheno2$Sex),perPheno2$G,xlab="Group",ylab="G",trace.label="Sex", fixed=TRUE)

#2
# install.packages("Matrix")
# library(Matrix)
library(effects)
perPheno2$Group <- as.factor(perPheno2$Group)
perPheno2$Sex <- as.factor(perPheno2$Sex)
model.lm <- lm(formula= G~Group*Sex, data=perPheno2)
perPheno2$predicted=predict(model.lm)
plot(effect(term="Group:Sex",mod=model.lm,default.levels=4),multiline=TRUE)


#3
df <- with(perPheno2 , aggregate(G, list(Group=Group, Sex=Sex), mean))
df$se <- with(perPheno2 , aggregate(G, list(Group=Group, Sex=Sex), function(x) sd(x)/sqrt(10)))[,3]

opar <- theme_update(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.background = element_rect(colour = "black"))
gp <- ggplot(df, aes(x=Group, y=x, colour=Sex, group=Sex))
gp + geom_line(aes(linetype=Sex), size=.6) + 
     geom_point(aes(shape=Sex), size=3) + 
     geom_errorbar(aes(ymax=x+se, ymin=x-se), width=.1)
theme_set(opar)



#### SAMPLE Size



# subset.mean <- function(x,y,z,n) { mean(subset(x, y == "z")$a, na.rm = TRUE)}
# mean.HCF.G <- subset.mean(perPheno1, perPheno1$dxGenInt, "HC.F", perPheno1$G)

#  HC.F IBS.F  HC.M IBS.M 
#     5    18     6     7 

mean.sd <- matrix(NA, nrow = 4,ncol = 2)
row.names(mean.sd) <- names(table(perPheno1$dxGenInt))
colnames(mean.sd) <- c("Mean", "SD")

mean.sd[1,1] <- mean(subset(perPheno1, perPheno1$dxGenInt == "HC.F")$G, na.rm = TRUE)
mean.sd[1,2] <- sd(subset(perPheno1, perPheno1$dxGenInt == "HC.F")$G, na.rm = TRUE)

mean.sd[2,1] <- mean(subset(perPheno1, perPheno1$dxGenInt == "IBS.F")$G, na.rm = TRUE)
mean.sd[2,2] <- sd(subset(perPheno1, perPheno1$dxGenInt == "IBS.F")$G, na.rm = TRUE)

mean.sd[3,1] <- mean(subset(perPheno1, perPheno1$dxGenInt == "HC.M")$G, na.rm = TRUE)
mean.sd[3,2] <- sd(subset(perPheno1, perPheno1$dxGenInt == "HC.M")$G, na.rm = TRUE)

mean.sd[4,1] <- mean(subset(perPheno1, perPheno1$dxGenInt == "IBS.M")$G, na.rm = TRUE)
mean.sd[4,2] <- sd(subset(perPheno1, perPheno1$dxGenInt == "IBS.M")$G, na.rm = TRUE)

mean.sd
#    Mean        SD
# HC.F  68.45267  9.046390
# IBS.F 86.06213 25.926123
# HC.M  44.57064  7.610016
# IBS.M 71.97551 13.962806

mean.sd.t <- t(mean.sd)
mean.sd.t 

md1 <- function(x,y,z) {return(x[1,y]-x[1,z])} 
md1(mean.sd.t,1,2)
#-17.60
md1(mean.sd.t,2,3)
#41.49
md1(mean.sd.t,3,4)
#-27
md1(mean.sd.t,1,4)
# -3.5
md1(mean.sd.t,2,4)
# 14
md1(mean.sd.t,1,3)
# 23.88203

## sample size
# install.packages("samplesize")
library(samplesize)

n.ttest(power = 0.8, alpha = 0.05, mean.diff = 41.49, sd1 = 25.926123, sd2 = 7, design = "unpaired", variance="unequal")
# $`Total sample size`
# [1] 9
# 
# $`Sample size group 1`
# [1] 7
# 
# $`sample size group 2`
# [1] 2

n.ttest(power = 0.8, alpha = 0.05, mean.diff = 3.5, sd1 = 9.04639, sd2 = 13.96281, design = "unpaired", variance="unequal")
# $`Total sample size`
# [1] 343
# 
# $`Sample size group 1`
# [1] 135
# 
# $`sample size group 2`
# [1] 208

n.ttest(power = 0.8, alpha = 0.05, mean.diff = 17, sd1 = 9.04639, sd2 = 13.96281, design = "unpaired", variance="unequal")


# These are retrospective calculations of sample sizes which may not be appropriate to prove that this study was powered since it is based on the study in question, if we had to design a new study with knowledge of standard deviation of the population and expected mean difference, we would need atleast 18 samples for detecting the largest differnces or 600 for smallest differnces for each comparison.


# Since we will be looking at 6 combinations; 18*6 = 108

n.ttest(power = 0.8, alpha = 0.05, mean.diff = 17, sd1 =25, sd2 = 13.96281, design = "unpaired", variance="unequal")

```

















#how many common samples between microbiome and permeability

```{r, echo=FALSE,include=FALSE}
load("C:/Users/swapnajoshi-admin/Documents/swapna/genus1.rda")
comSamples <- per[row.names(per)%in%substr(colnames(genus1),8,12),]
table(comSamples$dx) # HC IBS 4  10
table(comSamples$bowel.habits) # C D M N 3 3 3 5 
table(comSamples$dx, comSamples$Sex)
#       F M
#   HC  1 3
#   IBS 8 2
```

```{r, echo=FALSE,include=FALSE}
load("C:/Users/swapnajoshi-admin/Documents/swapna/physeq.rda")
load("C:/Users/swapnajoshi-admin/Documents/GIT_workspace/InetstinalPermeability/per.rda")

```

```{r, echo=FALSE,include=FALSE}
ex3 <- subset_samples(physeq, substr(Description,8,12)%in%row.names(per))
ex3
```


```{r, echo=FALSE,include=FALSE}
comSamples1 <- comSamples[ substr(row.names(sample_data(ex3)),8,12), ]
sample_data(ex3)$per_G <- comSamples1$highPerm

```
## Differences in richness between disease groups in the small sample
```{r, echo=FALSE,include=FALSE}
sample_variables(ex3)[1:10]

p <- plot_richness(ex3, x = "as.factor(Group)")
q <- p + geom_boxplot(data = p$data, aes(x = as.factor(Group), y = value, color = NULL), alpha = 0.1)
ggsave(q, file ="richness_IBSvsHCs.png")
q
```
## Differences in richness between bowel habit subtypes
```{r, echo=FALSE,include=FALSE}
p <- plot_richness(ex3, x = "as.factor(BH)")
q <- p + geom_boxplot(data = p$data, aes(x = as.factor(BH), y = value, color = NULL), alpha = 0.1)
ggsave(q, file ="richness_bowelhabits.png")
q
```
## Differences in richness between permeability G classes
```{r, echo=FALSE,include=FALSE}
p <- plot_richness(ex3, x = "as.factor(per_G)")
q <- p + geom_boxplot(data = p$data, aes(x = as.factor(per_G), y = value, color = NULL), alpha = 0.1)
ggsave(q, file ="richness_bowelhabits.png")
q
```
## Differences in richness between sex
```{r, echo=FALSE,include=FALSE}
p <- plot_richness(ex3, x = "as.factor(Gender)")
q <- p + geom_boxplot(data = p$data, aes(x = as.factor(Gender), y = value, color = NULL), alpha = 0.1)
ggsave(q, file ="richness_Sex.png")
q
```

## Differences in richness interaction between IBS and Sex
```{r, echo=FALSE,include=FALSE}
p <- plot_richness(ex3, x = "Dx_Gender_interaction")
q <- p + geom_boxplot(data = p$data, aes(x = as.factor(Dx_Gender_interaction), y = value, color = NULL), alpha = 0.1)
ggsave(q, file ="richness_interaction_Dx_sex.png")
q
```

# Abundance bar plots
```{r, echo=FALSE,include=FALSE}
TopNOTUs = names(sort(taxa_sums(ex3), TRUE)[1:20])
ent10 = prune_taxa(TopNOTUs, ex3)
png("BarPlot_top10_Group.png", height = 3000, width = 3500, res = 200)
plot_bar(ent10, "Group", fill = "Genus",facet_grid = ~Genus)
dev.off()
plot_bar(ent10, "Group", fill = "Family",facet_grid = ~Family)

```


# Abundance bar plots for permeability
```{r, echo=FALSE,include=FALSE}
TopNOTUs = names(sort(taxa_sums(ex3), TRUE)[1:20])
ent10 = prune_taxa(TopNOTUs, ex3)
png("BarPlot_top10_Group.png", height = 3000, width = 3500, res = 200)
plot_bar(ent10, "as.factor(per_G)", fill = "Genus",facet_grid = ~Genus)
dev.off()
plot_bar(ent10, "as.factor(per_G)", fill = "Family",facet_grid = ~Family)

```
# For clinical traits table 
```{r}
hcPer <- perMap[perMap$Group==1,];dim(hcPer)
mean(hcPer$HAD_Anxiety)

```

# Calculating the 'Reference range' for 'G' :
```{r}

# Reference Range = mean+2SD 

```
# unused edge R from Permeability5262017.Rmd file
# EdgeR

```{r}
design           <- model.matrix(~0+colDatPerm$IBS_Sex_int)
colnames(design) <- c("HC.F","IBS.F","HC.M", "IBS.M") #levels(colDatPerm$IBS_Sex_int)
```


```{r}
library(edgeR)
library(org.Hs.eg.db)
y       <- DGEList(counts = exprColDatPerm, group = colDatPerm$IBS_Sex_int)
y$Genes <- mapIds(org.Hs.eg.db, keys=rownames(y), keytype="ENSEMBL", column="SYMBOL", multiVals="first")
y       <- calcNormFactors(y)
y       <- estimateDisp(y,design, robust=TRUE)
```

```{r}
fit  <- glmQLFit(y, design, robust=TRUE)
F_M  <- makeContrasts(F-M, levels=design)
res1 <- glmQLFTest(fit, contrast=F_M)
res2 <- glmQLFTest(fit, contrast=F_M)
res3 <- glmQLFTest(fit, contrast=F_M)
res4 <- glmQLFTest(fit, contrast=F_M)

resultsFVsM <- data.frame(topTags(res,n=dim(res$table)[1]))
resultsFVsM <- data.frame(topTags(res,n=dim(res$table)[1]))
resultsFVsM <- data.frame(topTags(res,n=dim(res$table)[1]))
resultsFVsM <- data.frame(topTags(res,n=dim(res$table)[1]))

y$Genes <- y$Genes[row.names(resultsFVsM )]
resultsFVsM $Gene_Symbol <- y$Genes
```

## Association with Bowel Habits

```{r}
### BH
colDatPermCD <- subset(colDatPerm, colDatPerm$BH_Colon_Exam=="C"|colDatPerm$BH_Colon_Exam=="D"); dim(colDatPermCD)
# [1] 30  24
exprColDatPermCD <- exprColDat[,colnames(exprColDat)%in%row.names(colDatPermCD)]; dim(exprColDatPermCD)
# [1] 60204   30
all.equal(colnames(exprColDatPermCD),row.names(colDatPermCD))
colDatPermCD$BH_Colon_Exam <- factor(colDatPermCD$BH_Colon_Exam)
colDatPermCD$Lane <- factor(colDatPermCD$Lane)
QBHCDRes <- deseqOut(exprColDatPermCD, colDatPermCD, "BH_Colon_Exam", "Lane", 5, featureData)
QBHCDResFdr01 <- as.data.frame(subset(QBHCDRes, QBHCDRes$padj<0.1)); dim(QBHCDResFdr01)
QBHCDResTjp <- QBHCDRes[QBHCDRes$V2%in%tjp,c(28,34,31)]
###
colDatPermCN <- subset(colDatPerm, colDatPerm$BH_Colon_Exam=="C"|colDatPerm$BH_Colon_Exam=="N"); dim(colDatPermCN)
# [1] 30  24
exprColDatPermCN <- exprColDat[,colnames(exprColDat)%in%row.names(colDatPermCN)]; dim(exprColDatPermCN)
# [1] 60204   30
all.equal(colnames(exprColDatPermCN),row.names(colDatPermCN))
relevel()
colDatPermCN$BH_Colon_Exam <- factor(colDatPermCN$BH_Colon_Exam)
colDatPermCN$BH_Colon_Exam <- relevel(colDatPermCN$BH_Colon_Exam,ref="N")
colDatPermCN$Lane <- as.factor(colDatPermCN$Lane)
QBHCNRes <- deseqOut(exprColDatPermCN, colDatPermCN, "BH_Colon_Exam", "Lane", 5, featureData)
QBHCNResFdr01 <- as.data.frame(subset(QBHCNRes, QBHCNRes$padj<0.1)); dim(QBHCNResFdr01)
QBHCNResTjp <- QBHCNRes[QBHCNRes$V2%in%tjp,c(28,34,31)]

###
colDatPermDN <- subset(colDatPerm, colDatPerm$BH_Colon_Exam=="D"|colDatPerm$BH_Colon_Exam=="N"); dim(colDatPermDN)
# [1] 38  24
exprColDatPermDN <- exprColDat[,colnames(exprColDat)%in%row.names(colDatPermDN)]; dim(exprColDatPermDN)
# [1] 60204   38
all.equal(colnames(exprColDatPermDN),row.names(colDatPermDN))
colDatPermDN$BH_Colon_Exam <- factor(colDatPermDN$BH_Colon_Exam)
colDatPermDN$BH_Colon_Exam <- relevel(colDatPermDN$BH_Colon_Exam,ref="N")
colDatPermDN$Lane <- as.factor(colDatPermDN$Lane)
QBHDNRes <- deseqOut(exprColDatPermDN, colDatPermDN, "BH_Colon_Exam", "Lane", 5, featureData)
QBHDNResFdr01 <- as.data.frame(subset(QBHDNRes, QBHDNRes$padj<0.1)); dim(QBHDNResFdr01)
QBHDNResTjp <- QBHDNRes[QBHDNRes$V2%in%tjp,c(28,34,31)]

```